---
title: "MeniscusTrackingRefinement"
author: "Tomas Pierce"
date: "2022-10-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries
```{r lib load, include=FALSE}
require(ggplot2)
require(dplyr)
require(psych)
require(sys)
require(purrr)
require(gridExtra)
```

Load in data
```{r data load, include=FALSE}
setwd("/home/tomas/Projects/DrinkMoth2")
names <- Sys.glob("chopped_men_data/*.tsv")
readtsv <- function (name) {
  read.csv(name, header=FALSE, sep="\t")
}
tables <- lapply(names, readtsv)
for (i in 1:length(names)) {
  tables[[i]]$Filename <- names[[i]]
}
```

rename key columns
```{r renamin,g, include=FALSE}
recol <- function (table) {
  colnames(table)[1] <- "Frame"
  colnames(table)[2] <- "Row"
  colnames(table)[3] <- "Col"
  colnames(table)[4] <- "Dcenter"
  colnames(table)[5] <- "Area"
  colnames(table)[6] <- "AreaBbox"
  colnames(table)[7] <- "AreaConvex"
  colnames(table)[8] <- "AreaFilled"
  colnames(table)[9] <- "AxisMajorLen"
  colnames(table)[10] <- "AxisMinorLen"
  colnames(table)[11] <- "Eccentricity"
  colnames(table)[12] <- "EquivDiamArea"
  colnames(table)[13] <- "EulerNumber"
  colnames(table)[14] <- "Extent"
  colnames(table)[15] <- "FeretDiameterMax"
  colnames(table)[16] <- "IntensityMax"
  colnames(table)[17] <- "IntensityMean"
  colnames(table)[18] <- "IntensityMin"
  colnames(table)[19] <- "Label"
  colnames(table)[20] <- "Orientation"
  colnames(table)[21] <- "Perimeter"
  colnames(table)[22] <- "PerimeterCrofton"
  colnames(table)[23] <- "Solidity"
  colnames(table)[24] <- "CenterY"
  colnames(table)[25] <- "CenterX"
  table
}

named_tables <- lapply(tables,recol)
```

Filter to only videos in which we have collected at least 10 data points (rows)
```{r Give me the good data, include=FALSE}
has10 <- function(table) {
  nrow(table) > 10
}

long_tables <- named_tables[sapply(named_tables, has10) == TRUE]
```

We can see the trend line in the data we extract, we just want to
see if there's a way to refine it further
```{r show data, include=TRUE}
graphs <- list()
for (longs in long_tables) {
  p <- ggplot(longs, aes(Frame, Row, col=Perimeter)) + geom_point() +
    theme_bw() + labs(title = longs$Filename[1])# + ylim(0, 600)
  show(p)
  graphs[[length(graphs) + 1]] <- p
  # longs$Wideness <- cos(longs$Orientation) * longs$MajorAxisLen
}

```
We can see that some of the videos go better than others.
We take 14 and check to see if its good, then use it as an example of
how we would want to do further analysis.
```{r checking stuff, include=TRUE}
twelve = long_tables[[12]]
ggplot(twelve, aes(Frame, Row, col=Extent)) + geom_point() + labs(
  title = twelve$Filename[1]
) + theme_bw()
```

As we can see that item 12 is one where the current approach works particualy well,
we can use it work out a method of determining a local derivative measurement
which is defensible from an approaches perspective.
```{r looking for ways to find local derivates, include=TRUE}
spline <- loess(Row ~ Frame, twelve)
predicted <- predict(spline)
ggplot(twelve, aes(Frame, Row)) + 
  geom_point(aes(col="Measurements")) + 
  geom_line(aes(Frame, predicted, col="Model")) + 
  theme_bw() + 
  labs(title = twelve$Filename[1]) +
  theme(legend.position = c(0.2, 0.75))

act_x_difs <- diff(twelve$Frame)
act_y_difs <- diff(twelve$Row)
pred_y_difs <- diff(predicted)

from_act_deriv <- act_y_difs / act_x_difs
from_mod_deriv <- pred_y_difs / act_x_difs

derivs = twelve[1:nrow(twelve)-1,]

derivs$NaiveDerive <- from_act_deriv
derivs$LoessDerive <- from_mod_deriv

ggplot(derivs, aes(Frame, NaiveDerive)) + 
  geom_point() +
  theme_bw() + 
  labs(title = twelve$Filename[1])

ggplot(derivs, aes(Frame, LoessDerive)) + 
  geom_point() +
  theme_bw() + 
  labs(title = twelve$Filename[1])
```

Generalize this to every single file
```{r, include=TRUE}
splines <- lapply(long_tables, function(table) {
  loess(Row ~ Frame, table)
})
predictions <- lapply(splines, function(spline) {
  predict(spline)
})
for (i in 1:length(long_tables)) {
  table <- long_tables[[i]]
  model_pred <- predictions[[i]]
  frame_difs <- diff(table$Frame)
  row_difs <- diff(model_pred)
  deriv <- row_difs / frame_difs
  derivs <- table[1:nrow(table)-1,]
  derivs$LoessDerive <- deriv
  p <- ggplot(derivs, aes(Frame, LoessDerive)) +
    geom_point() + 
    theme_bw() + 
    labs(title = derivs$Filename[1]) + 
    geom_hline(yintercept = 0, color='red')
  show(p)
}
```

Permit me to try filtration on Major Axis Length
```{r give me the glorious graphs, include=TRUE}

narrows <- lapply(long_tables, function (table) {
  table[table$AxisMajorLen < 35,]
})

for (nar in narrows) {
  filename <- nar$Filename[1]
  name_bits <- unlist(strsplit(filename, "/"))
  name <- name_bits[2]
  pieces <- unlist(strsplit(name, "[\\_\\-\\.]"))
  graph_title <- paste(pieces[-length(pieces)], collapse=' ')
  p<- ggplot(nar, aes(Frame / 30, Row)) + 
    geom_point() +
    theme_bw() + 
    labs(title = graph_title) +
    ylab("Meniscus Depth (pixels)") +
    xlab("Elapsed Time (seconds)")
  
  show(p)
}
```
Create Loess model derivatives on filters
```{r give me better derivatives, include=TRUE}
for (nar in narrows) {
  filename <- nar$Filename[1]
  name_bits <- unlist(strsplit(filename, "/"))
  name <- name_bits[2]
  pieces <- unlist(strsplit(name, "[\\_\\-\\.]"))
  graph_title <- paste(pieces[-length(pieces)], collapse=' ')
  graph_title <- paste(graph_title, "Meniscus Rate")
  loess_model <- loess(Row ~ Frame, nar)
  preds <- predict(loess_model)
  filename <- nar$Filename[1]
  delta_x <- diff(nar$Frame)
  delta_y <- diff(preds)
  deriv <- delta_y / delta_x
  dtable <- nar[1:nrow(nar) - 1,]
  dtable$Deriv <- deriv
  # convert unit from pixels / frame to pixels / second
  dtable$MenRate <- deriv * 30
  # convert time to seconds
  dtable$ElapsedTime <- dtable$Frame / 30
  p <- ggplot(dtable, aes(ElapsedTime, MenRate)) + 
    geom_point() + 
    theme_bw() +
    labs(title = graph_title) + 
    geom_hline(yintercept = 0, color='red') + 
    ylab("Meniscus Velocity (pixels / second)") + 
    xlab("Elapsed Time (seconds)")
  show(p)
}
```